---
title: "p8105_hw1_rk3445"
author: "Rosie K"
date: "2025-09-16"
output: github_document
---

## Problem 1

```{r echo = TRUE, message = FALSE}
library(tidyverse)
```

```{r}
library(moderndive)
data("early_january_weather")
```

The `early_january_weather` dataset has `r nrow(early_january_weather)` observations and `r ncol(early_january_weather)` variables. 

The variables include:

- `origin`: the origin of the observation
- `year`,`month`,`day`,`hour` : Time of recording.
- `temp`: temperature in Fahrenheit
- `dewp` : dewpoint in Fahrenheit
- `humidity`: humidity 
- `wind_dir`,`wind_speed`,`wind_gust`: wind direction/speed/gust speed
- `precip`: precipitation
- `pressure`: pressure
- `visib`: visibility

The temperature (`temp`) ranges from `r min(early_january_weather$temp)` to `r max(early_january_weather$temp)` degrees Fahrenheit, with a mean of `r round(mean(early_january_weather$temp),1)`. 


```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp)) + geom_point(aes(color = humid)) +labs(
    title = "Temperature in early January",
    x = "Day",
    y = "Temperature (°F)"
  )
```

The scatter plot shows the temperature trends in January. The temperature increased steadily over two weeks, peaking around mid-January above 55°F. As temperature increased, humidity also varied, with spikes of high humidity during the warmest days. Early January started cooler and drier, but by mid-January it became warmer and more humid.

```{r echo=FALSE,message = FALSE}
ggsave("temperature.pdf")
```


## Problem 2


Create a data frame comprised of:

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r}
df = 
  tibble(
    numeric_var = rnorm(10),
    logical_var = numeric_var > 0,
    char_var = letters[1:10],
    factor_var = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
  )
```


**Try to take the mean of each variable in your dataframe. What works and what doesn’t?**

* Method 1) 
```{r}
sapply(df, mean)
```



* Method 2)
Hint: for now, to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.

```{r}
num <- df %>% pull(numeric_var)
mean(num)

logi <- df %>% pull(logical_var)
mean(logi)

char <- df %>% pull(char_var)
mean(char)

fac <- df %>% pull(factor_var)
mean(fac)
     
```

Numeric vector can calculated mean. R treats TRUE as 1 and FALSE as 0 for logical vector, so mean gives the proportion of TRUEs. However, character vector is text so mean cannot be calculated. Factor vector is categorical value so mean cannot be calculated as well.



**In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?**

```{r echo = TRUE, message=FALSE}
num_char <- mean(as.numeric(char))
num_log <- mean(as.numeric(logi))
num_fac <- mean(as.numeric(fac))
```

For character variables, coercing them to numeric produces a warning because non-numeric strings cannot be converted, resulting in NA values and an invalid mean. Converting logical values into numeric values gives the proportion of TRUE values. For factor variables, numeric conversion returns the internal integer codes assigned to each level. However, taking the mean of these integer codes is not related to actual mean of the values.
This helps me to explain why mean() function behaves differently depending on the variable type and it only works directly on numeric and logical vectors.


